\section{References}
\markboth{References}{References}
\thispagestyle{myheadings}
\label{references}

\begin{thebibliography}{99}

\bibitem{cci1}
CCITT: {\em Red Book} Volume VIII -- Fascicle VIII.7,
Data Communication Networks:
{\em Message Handling Systems,
Recommendations X.400-X.430 (1984)}.
Malaga-Torremolinos, 1984.

\bibitem{cci2}
CCITT: {\em Blue Book} Volume VIII -- Fascicle VIII.7,
Data Communication Networks:
{\em Message Handling Systems,
Recommendations X.400-X.420 (1988)}.
Geneva, 1989.
(see also: ISO 10021)

\bibitem{cci3}
CCITT: {\em Blue Book} Volume VIII -- Fascicle VIII.8,
Data Communication Networks:
{\em Directory, Recommendations X.500-X.521 (1988)}.
Geneva, 1989.
(see also: ISO 9594)

\bibitem{cci4}
CCITT: Recommendation X.509.
{\em The Directory -- Authentication Framework}.
Geneva, 1989.
(see also: ISO 9594-8)

\bibitem{iso8}
ISO-6523:
Information Processing Systems --
Data Interchange --
{\em Structure for the Identification of Organizations}.
1984.

\bibitem{iso1}
ISO 7489:
Information Processing Systems --
{\em Open Systems Interconnection} --
{\em Basic Reference Model}.

\bibitem{iso2}
ISO 7489-2:
Information Processing Systems --
{\em Open Systems Interconnection} --
Basic Reference Model.
Part 2: {\em Security Architecture}.

\bibitem{iso7}
ISO-8824:
Information Processing Systems --
Open Systems Interconnection --
{\em Specification of Abstract Syntax Notation One (ASN.1)}.
Second Edition 1990.\\
ISO-8825:
Information Processing Systems --
Open Systems Interconnection --
{\em Specification of Basic Encoding Rules of
Abstract Syntax Notation One (ASN.1)}.
Second Edition 1990.

\bibitem{iso4}
ISO 9594:
Information Processing Systems --
Open Systems Interconnection --
{\em The Directory}.
(see also: CCITT-X.500 Recommendations)

\bibitem{iso5}
ISO 9735:
{\em Electronic Data Interchange For Administration, Commerce
and Transport (EDIFACT) -- Application Level Syntax Rules}.
1988.

\bibitem{iso6}
ISO-10021:
Information Processing Systems --
Text Communication --
{\em MOTIS}.
(see also: CCITT-X.400-1988 Recommendations)

\bibitem{iso3}
ISO 7816:
{\em Identification Cards --
Integrated Circuit(s) Cards with Contacts}.
1989.

\bibitem{ecm1}
ECMA:
{\em Security in Open Systems}.
\\
1. ECMA TR/46:
{\em A Security Framework}.
July 1988, 71 pages.
\\
2. Standard ECMA-138:
{\em Data Elements and Service Definitions}.
December 1989, 81 pages.

\bibitem{rfc1}
{\em Privacy Enhancement for Internet Electronic Mail}.
Internet Engineering Task Force (IETF) \\
RFC 1421. John Linn.  Privacy Enhancement for Internet Electronic Mail:  Part I:
Message Encryption and Authentication Procedures.  Internet Proposed Standard,
1993. \\
RFC 1422. Stephen Kent.  Privacy Enhancement for Internet Electronic Mail:  Part
II: Certificate-Based Key Management.  Internet Proposed Standard, 1993 \\
RFC 1423. David Balenson.  Privacy Enhancement for Internet Electronic Mail:
Part III: Algorithms, Modes, and Identifiers.  Internet Proposed Standard, 1993 \\
RFC 1424. B. Kaliski.  Privacy Enhancement for Internet Electronic Mail:  Part IV:
Key Certification and Related Services.  Internet Proposed Standard, 1993

\bibitem{Stable}
NIST-OIW: {\em Stable Implementation Agreements for Open Systems}, 
December 1992

\bibitem{Working}
NIST-OIW: {\em Working Implementation Agreements for Open Systems}, 
December 1991

\bibitem{TeleTrusT}
TeleTrust: 1992, {\em Object Identifiers -- Register}

\bibitem{riv1}
R. Rivest, A. Shamir, L. Adleman:
{\em A Method  for Obtaining Digital Signatures
and Public Key Cryptosystems}.
Communications of the ACM, 21(2), 120-126, Feb 1978.

\bibitem{ElGamal}
T. ElGamal:  {\em A  public key  cryptosystem  and  a signature
scheme  based  on discrete logarithms}, IEEE  Transactions on
Information    Theory,  IT-31,  Number  4,  July  1985,  pp.
469-472.

\bibitem{NistElGamal}
ElGamal-NIST-OIW: {\em Stable Implementation Agreements for Open Systems}, 
Part 12, Annex D, December 1991

\bibitem{nis1}
Federal Information  Processing Standards Publication  (FIPS
PUB)   46-1, {\em Data  Encryption Standard},  U.S. Department  of
Commerce/National Bureau  of Standards, Supersedes  FIPS PUB
46,  January 15, 1977, Reaffirmed January 22, 1988. \\ [1ex]
ANSI   X3.92-1981,  Data   Encryption  Algorithm,   American
National  Standards Institute, Approved December 30, 1980. \\ [1ex]
Federal Information  Processing Standards  Publication (FIPS
PUB)    81,  DES  Modes  of Operation,  U.S.  Department  of
Commerce/National Bureau of Standards, December 2, 1980. \\ [1ex]
ANSI  X3.106-1983,  Data  Encryption  Algorithm -  Modes  of
Operation, American  National Standards  Institute, Approved
May  16, 1983. \\ [1ex]
Federal Information  Processing Standards  Publication (FIPS
PUB)  74, Guidelines for Implementing and Using the NBS Data
Encryption  Standard, U.S.  Department of  Commerce/National
Bureau of Standards, April 1, 1981.

\bibitem{MD2}
Burt Kaliski:
The MD2 Message-Digest Algorithm.
RFC 1319 [MD2], RSA Laboratories, April 1992, 17 pages.

\bibitem{MD4}
Ronald Rivest:
The MD4 Message-Digest Algorithm.
RFC 1320 [MD4], RSA Data Security Inc., April 1992, 20 pages.
(Obsoletes RFC 1186.)

\bibitem{MD5}
Ronald. Rivest:
The MD5 Message-Digest Algorithm.
RFC 1321 [MD5], RSA Data Security Inc., April 1992, 21 pages.

\bibitem{cha1}
D. Chaum:
{\em Sicherheit ohne Identifizierung}.
Informatik-Spektrum, 10, 262-277, 1987.
(Receivable from the author:
{\em Security without Identification:
Card Computers to Make Big Brother Obsolete}.
David Chaum, 1987, 16 pages.)

\bibitem{gri2}
R. Grimm:
{\em Security on Networks -- Do We Really Need It?}
Computer Networks and ISDN Systems,
Volume 17 (1989), Numbers 4 \& 5, 315-323, October 1989.
Elsevier Science Publishers B.V. (North-Holland), 1989.

\bibitem{gri3}
R. Grimm:
{\em Security for Open Communication --
The DFN Mail Security Project Report}.
Computer Networks and ISDN Systems,
Volume 19 (1990), Numbers 3-5, 196-200, November 1990.
Elsevier Science Publishers B.V. (North-Holland), 1990.

\bibitem{gri4}
R. Grimm, R. Nausester, W. Schneider, U. Viebeg:
{\em Secure DFN}.
1. Principles of Security Operations.
2. Specification of Security Interfaces.
Arbeitspapiere der GMD 491, ISSN 0723-05008.
GMD, Birlinghoven, November 1990, 167 pages.

\bibitem{gri5}
R. Grimm, D. Heagerty:
{\em Recommendation for a Shorthand X.400 Address Notation}.
Computer Networks and ISDN Systems,
Volume 17 (1989), Numbers 4 \& 5, 263-267, October 1989.
Elsevier Science Publishers B.V. (North-Holland), 1989.

\bibitem{mit}
C. Mitchell, D. Rush, M. Walker:
{\em A Secure Messaging Architecture Implementing the X.400-1988 Features}.
The Computer Journal, Vol 33, No. 4, 290-295, 1990.
Published on behalf of The British Computer Society
by Cambridge University Press.

\bibitem{ord1}
H.-J. Ordemann, R. Schomerus:
{\em Bundesdatenschutzgesetz}.
Gesetz zum Schutz vor Missbrauch personenbezogener Daten bei
der Datenverarbeitung.
Beck Verlag, M\"unchen, 1977, 204 pages.

\bibitem{qua1}
J.S. Quarterman, J.C. Hoskins:
{\em Notable Computer Networks}.
Communications of the ACM, 29(10), 932-971, 1986.

\bibitem{ISODE}
M. Rose: {\em The ISO Development Environment}, User's Manual, Version 8.0

\bibitem{wsc1}
W. Schneider:
{\em Sicherheit bei X.400-Systemen}.
Datenschutz und Datensicherung (DuD) 1/90, 32 ff.
Vieweg Verlag, Wiesbaden, 1990.

\bibitem{ste1}
J.G. Steiner, C. Neumann, J.I. Schiller:
{\em Kerberos: An Authentication Service for Open Network Systems}.
USENIX Winter Conference February 9-12, Dallas, Texas, 1988.
(Project Athena, MIT, Boston, MA).

\bibitem{str1}
B. Struif:
{\em Chipkarten und Normung}.
Datenschutz und Datensicherung (DuD) 1/90, 23-30.
Vieweg Verlag, Wiesbaden, 1990.

\bibitem{ttt1}
TeleTrusT:
{\em Open Shops for Informations Services (OSIS):
Die elektronische Unterschrift aus dem Mikrochip}.
GMD-Spiegel, 1/86, 6 articles, March 1986.
GMD, Birlinghoven, 1986.

\bibitem{ttt2}
TeleTrusT:
{\em Open Shops for Information Services (OSIS) Final Report}.
Community--COST,
Concertation Agreement on Concerted Action Research Project
in the Field of Teleinformatics,
COST Project 11, OSIS.
Juni 1986.

\bibitem{ttt3}
TeleTrusT:
{\em Vertrauensw\"urdige elektronische Kommunikation}.
GMD-Spiegel, 1/88, 3 articles, March 1988.
GMD, Birlinghoven, 1988.

\bibitem{ttt4}
TeleTrusT:
{\em SmartCards - eine neue Dimension in der Informationstechnik}.
GMD-Spiegel, 1/92, 10 articles, March 1992.
GMD, Birlinghoven, 1992.

\end{thebibliography}
